---
title: "StatApps Project"
author: "A. Renganathan & Y.Baldonado"
date: "2023-04-07"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#install.packages("utils")
#install.packages("survey")
#install.packages("dplyr")
#install.packages("tidyverse")
#install.packages("jtools")
#install.packages("dataMaid")
#install.packages("expss")



#remove CRAN version of RNHANES PACKAGE
library(utils)
#detach("package:RNHANES", unload = TRUE)

#install devtools for the following
#install.packages("devtools")
library(devtools)
#install RNHANES updated vers from github
#install_github("silentspringinstitute/RNHANES")

#LOAD LIBRARIES
library(RNHANES) #package used to download NHANES data directly from the website
library(survey)#package used for complex survey design and analysis, incorporates weights
library(dplyr) #package used for recoding
library(tidyverse) #package used for recoding
library(MASS) #package for stepwise regression
library(jtools) #package for summarizing logistic regressions
library(ggplot2) #package for graphs
library(ggeffects)
library(pROC)


#NOTE ON DPLYR AND CAR
#recoding is done utilizing dplyr, detach car package to avoid errors
#if car is not loaded, the following will spit out an error
#detach(package:car, unload = TRUE)

```

#Prologue

##Purpose

##DATA

```{r data management}

#overall dataframe download for 2017-2018
nhanes<- nhanes_load_data(c("DEMO", "DR1TOT", "DSQTOT", "OSQ", "DBQ"), c("2017-2018","2017-2018","2017-2018","2017-2018","2017-2018"), cache = "./nhanes_data")


demo <- nhanes$DEMO %>%
  select(SEQN, #identifier
         SDDSRVYR,#survey year
         SDMVPSU, #masked variance pseudo-primary sampling units
         SDMVSTRA,  #masked variance pseudo-stratum
         WTMEC2YR,
         RIDAGEYR, RIAGENDR, RIDRETH3) %>%
  rename(age = RIDAGEYR) %>%
  rename(gender = RIAGENDR) %>%
  rename(race_ethnicity = RIDRETH3)

dietary<- nhanes$DR1TOT %>%
  select (SEQN, DR1TCALC, DR1TVD,
            WTDRD1) %>% #full sample 2 year mobile examination center(where day 1 diet interview occurred) weight) 
  rename(calcium_intake = DR1TCALC) %>%
  rename(vitamin_d= DR1TVD)

supplements <- nhanes$DSQTOT %>%
  select(SEQN, DSQTVD, DSQTCALC)%>%
  rename(vitamind_supp = DSQTVD) %>%
  rename(calcium_supp= DSQTCALC)

osteo <- nhanes$OSQ %>%
  select(SEQN, OSQ060, OSQ010B) %>%
  rename(osteoporosis = OSQ060) %>%
  rename(fractured_wrist = OSQ010B) 

milk <- nhanes$DBQ %>%
  select(SEQN, DBQ223A, DBQ223B, DBQ223C, DBQ223D, DBQ223E) %>%
  rename(regular_milk = DBQ223A) %>%
  rename(two_percent = DBQ223B) %>%
  rename(one_percent = DBQ223C) %>%
  rename(fatfree_skim = DBQ223D) %>%
  rename(soy_milk = DBQ223E) 
  
#remove larger nhanes file
rm(nhanes)


#joining dataframes by main identifier (SEQN) and limiting population to 50+yo due to OSQ data
osteo_data2 <- demo %>%
  full_join(dietary, by = "SEQN") %>%
  full_join(supplements, by = "SEQN") %>%
  full_join(osteo, by = "SEQN") %>%
  full_join(milk, by = "SEQN") %>%
  filter(age>=50)

#creating dairy variable to capture whether dairy milk is consumed = 1
osteo_data3 <- osteo_data2 %>% 
  mutate(dairy= ifelse(osteo_data2$regular_milk == 10 | osteo_data2$two_percent == 11 |osteo_data2$one_percent == 12 | osteo_data2$fatfree_skim == 13 | osteo_data2$soy_milk != 14, 1, 0))%>% 
  mutate(dairy = coalesce(dairy,0))#if dairy is NA then let = 0, otherwise keep value of 1 determined above 


#repeat but with nondairy
osteo_data4 <- osteo_data3 %>%
  mutate(nondairy = ifelse(osteo_data2$regular_milk != 10 | osteo_data2$two_percent != 11 |osteo_data2$one_percent != 12 | osteo_data2$fatfree_skim != 13 | osteo_data2$soy_milk == 14, 1, 0 ))%>%
  mutate(nondairy = coalesce(nondairy, 0))

#new dataset with newly created dairy variables (removing each milk option)
os <- osteo_data4 %>% select(age,SDDSRVYR, SDMVPSU,  SDMVSTRA,WTDRD1, WTMEC2YR, gender, race_ethnicity, calcium_intake, vitamin_d, vitamind_supp, calcium_supp,osteoporosis, fractured_wrist, dairy, nondairy)


count(os) 

#n=3069

```

```{r recoding data}
os.recoded<- os %>%
  mutate(race_ethnicity = recode_factor(.x=race_ethnicity,
                                  `1`= 1,
                                  `2`='Other Hispanic',
                                  `3`= 'White',
                                  `4`='African American',
                                  `6`='Asian',
                                  `7` = 'Mixed, other'))%>%
  mutate(gender= recode_factor(.x=gender,
                                 `1`='Male',
                                 `2`='Female')) %>%
  mutate(osteoporosis= recode_factor(.x=osteoporosis,
                                 `1`='Yes',
                                 `2`= 'No',
                                 `7`= NA_character_,
                                 `9`= NA_character_))%>%
  mutate(fractured_wrist= recode_factor(.x=fractured_wrist,
                                 `1`='Yes',
                                 `2`= 'No',
                                 `7`= NA_character_,
                                 `9`= NA_character_))%>%
  mutate(WTDRD1 = as.numeric(WTDRD1)) %>%
  mutate(WTMEC2YR = as.numeric(WTMEC2YR)) %>%
  mutate(WTDRD1 = coalesce(WTDRD1, WTMEC2YR))


rm(osteo_data2)
rm(osteo_data3)
rm(osteo_data4)
summary(os.recoded)
```

##Survey Design

```{r survey design}
#sample weight for aggregate data is INT2YR
#do not na.omit with survey data. you need the whole set for the weights to work
svydsgn <- svydesign(data = os.recoded,
                     id=~SDMVPSU,
                     strata = ~SDMVSTRA,
                     weights = ~WTDRD1,
                     nest=TRUE)

#subset by age
restr.age <- subset(svydsgn,
                        age>=50,
                        options(na.action = "na.pass"))



#subset by osteoporosis
svydsgn.osteo <- subset(restr.age,
                      osteoporosis=="Yes",
                      options(na.action="na.pass"))
svydsgn.no.osteo<-subset(restr.age,
                      osteoporosis=="No",
                   options(na.action="na.pass"))

#relevelling for logregs
restr.age$variables$osteoporosis<- relevel(restr.age$variables$osteoporosis, ref = "No")
restr.age$variables$race_ethnicity<- relevel(restr.age$variables$race_ethnicity, ref = "White")
restr.age$variables$fractured_wrist<- relevel(restr.age$variables$fractured_wrist, ref = "No")



osteo.svydsgn <- restr.age
```

```{r descr stats and chi-sq, echo=FALSE}
#age
svymean(~age, svydsgn.no.osteo)
svymean(~age, svydsgn.osteo)

#calculate standard deviation, called 'variance' below 
#you can only use this on numeric variables, not factors/categorical
sqrt(svyvar(~age, svydsgn.osteo))
sqrt(svyvar(~age, svydsgn.no.osteo))

#gender
svymean(~gender, svydsgn.no.osteo)*100
svymean(~gender, svydsgn.osteo)*100

#race eth
svymean(~race_ethnicity, svydsgn.no.osteo)*100
svymean(~race_ethnicity, svydsgn.osteo)*100

#calcium intake
svymean(~calcium_intake, svydsgn.osteo, na.rm = TRUE) 
svymean(~calcium_intake, svydsgn.no.osteo, na.rm = TRUE)

#vitamin d intake
svymean(~vitamin_d, svydsgn.osteo, na.rm = TRUE)
svymean(~vitamin_d, svydsgn.no.osteo, na.rm = TRUE) 

#calcium supp
svymean(~calcium_supp, svydsgn.osteo, na.rm = TRUE)
svymean(~calcium_supp, svydsgn.no.osteo, na.rm = TRUE)

#vitamin d supp
svymean(~vitamind_supp, svydsgn.osteo, na.rm = TRUE)
svymean(~vitamind_supp, svydsgn.no.osteo, na.rm = TRUE)

#fractured wrist
svymean(~fractured_wrist, svydsgn.osteo)*100
svymean(~fractured_wrist, svydsgn.no.osteo, na.rm = TRUE)*100

#dairy
svymean(~dairy, svydsgn.osteo)*100
svymean(~dairy, svydsgn.no.osteo)*100

#nondairy
svymean(~nondairy, svydsgn.osteo)*100
svymean(~nondairy, svydsgn.no.osteo)*100


#chi-sq tests for diff btwn osteoporosis status
svychisq(~age+osteoporosis ,design = osteo.svydsgn, statistic=c("Chisq")) #signf <0.0001
svychisq(~race_ethnicity+osteoporosis ,design = osteo.svydsgn, statistic=c("Chisq")) #not signf
svychisq(~gender+osteoporosis ,design = osteo.svydsgn, statistic=c("Chisq")) # signf <0.0001
svyttest(calcium_intake~osteoporosis, design=osteo.svydsgn) #signf <0.05
svyttest(vitamin_d~osteoporosis, design=osteo.svydsgn) #not signf
svyttest(vitamind_supp~osteoporosis, design=osteo.svydsgn)#not signf
svyttest(calcium_supp~osteoporosis, design=osteo.svydsgn) #signf <0.001
svychisq(~fractured_wrist+osteoporosis ,design = osteo.svydsgn, statistic=c("Chisq")) #not signf
svychisq(~dairy+osteoporosis ,design = osteo.svydsgn, statistic=c("Chisq")) #not signf
svychisq(~nondairy+osteoporosis ,design = osteo.svydsgn, statistic=c("Chisq")) #not signf


```




```{r univariate logreg models}

#AGE: significant
o1 <- svyglm(osteoporosis~ age, design = restr.age, family = binomial(link = "logit"), data = os.recoded)
summary(o1)

#GENDER: FEMALE significant
o2<- svyglm(osteoporosis~ gender, design = restr.age, family = binomial(link = "logit"), data = os.recoded)
summary(o2)

#VITAMIN D: not sig.
o3<- svyglm(osteoporosis~ vitamin_d, design = restr.age, family = binomial(link = "logit"), data = os.recoded)
summary(o3)

#CALCIUM INTAKE: significant
o4<- svyglm(osteoporosis~ calcium_intake, design = restr.age, family = binomial(link = "logit"), data = os.recoded)
summary(o4)

#VITAMIN D SUPP: not sig
o5<- svyglm(osteoporosis~ vitamind_supp, design = restr.age, family = binomial(link = "logit"), data = os.recoded)
summary(o5)

#CALCIUM SUP:very significant
o6<- svyglm(osteoporosis~ calcium_supp, design = restr.age, family = binomial(link = "logit"), data = os.recoded)
summary(o6)

#FRACTURED WRIST: not sig.
o7<- svyglm(osteoporosis~ fractured_wrist, design = restr.age, family = binomial(link = "logit"), data = os.recoded)
summary(o7)

#NONDAIRY: not sig
o8<- svyglm(osteoporosis~ nondairy, design = restr.age, family = binomial(link = "logit"), data = os.recoded)
summary(o8)

#DAIRY: not sig
o9<- svyglm(osteoporosis~ dairy, design = restr.age, family = binomial(link = "logit"), data = os.recoded)
summary(o9)
```

```{r multivariate logreg models}

##### significant predictors are: age, gender, calcium intake, and calcium supplementation######

#osteo with all predictors
osteo_model0<- svyglm(osteoporosis ~ age + race_ethnicity + gender + vitamin_d + vitamind_supp + calcium_intake + calcium_supp + fractured_wrist + nondairy + dairy, design = restr.age, family = binomial(link= "logit"), data = os.recoded)

summ(osteo_model0,
     exp=TRUE,
     confint=TRUE,
     model.fit = TRUE,
     model.info = TRUE,
     digits=2,
     vifs = TRUE)
summary(osteo_model0) #,correlation = FALSE, df.resid = NULL)

#basic predictors
osteo_basic<- svyglm(osteoporosis ~ age + race_ethnicity + gender, design = restr.age, family = binomial(link= "logit"), data = os.recoded)
summ(osteo_basic,
     exp=TRUE,
     confint=TRUE,
     model.fit = TRUE,
     model.info = TRUE,
     digits=2,
     vifs = TRUE)
summary(osteo_basic)

#basic + milk: milks have no effect
osteo_milk<- svyglm(osteoporosis ~ age + race_ethnicity + gender +  nondairy + dairy , design = restr.age, family = binomial(link= "logit"), data = os.recoded)

summ(osteo_milk,
     exp=TRUE,
     confint=TRUE,
     model.fit = TRUE,
     model.info = TRUE,
     digits=2,
     vifs = TRUE)
summary(osteo_milk)

#basic + nutrients: vit D & calcium
osteo_nutrients<- svyglm(osteoporosis ~ age + race_ethnicity + gender + vitamin_d + vitamind_supp + calcium_intake + calcium_supp, design = restr.age,  family = binomial(link= "logit"), data = os.recoded)
summ(osteo_nutrients,
     exp=TRUE,
     confint=TRUE,
     model.fit = TRUE,
     model.info = TRUE,
     digits=2,
     vifs = TRUE)
summary(osteo_nutrients)


#basic + calcium:
osteo_calc<- svyglm(osteoporosis ~ age + race_ethnicity + gender + calcium_intake + calcium_supp, design = restr.age,  family = binomial(link= "logit"), data = os.recoded)
summ(osteo_calc,
     exp=TRUE,
     confint=TRUE,
     model.fit = TRUE,
     model.info = TRUE,
     digits=2,
     vifs = TRUE)
summary(osteo_calc)

#basic + vitamin D:
osteo_vitd<- svyglm(osteoporosis ~ age + race_ethnicity + gender + vitamin_d + vitamind_supp, design = restr.age,  family = binomial(link= "logit"), data = os.recoded)
summ(osteo_vitd,
     exp=TRUE,
     confint=TRUE,
     model.fit = TRUE,
     model.info = TRUE,
     digits=2,
     vifs = TRUE)
summary(osteo_vitd)





#model with only sig predictors:
osteo_sig <- svyglm(osteoporosis ~ age +  race_ethnicity + gender + calcium_supp, design = restr.age,  family = binomial(link= "logit"), data = os.recoded)

summ(osteo_sig,
     exp=TRUE,
     confint=TRUE,
     model.fit = TRUE,
     model.info = TRUE,
     digits=2,
     vifs = TRUE)
summary(osteo_sig)

#multicollinearity check
vif(osteo_sig) # all below 5, no issues with multicollinearity

#logreg curve for best model:
####logistic reg plot#####
# Model: osteo_sig
#need to have install.packages("ggeffects")
library(ggeffects)
plot(ggpredict(osteo_sig), type = "response",)



```

#Assumption Checking

1. Observations are independent. - MET
2. No Multicollinearity - NOT MET
      Evidence of multicollinearity evident when VIF scores exceeded 10 for continuous variables. For categorical variables, this was checked using chi-square tests.  This was especially present for INSERT RELEVANT VARIABLES IN RELEVANT MODELS HERE across the logistic regressions.
      
3. Linearity - MET
    The only continuous variable in the logistic regressions is age, so we test its linearity across all the models. All logistic regressions met the assumption of linearity for age. 

```{r assumption checking}


```
